---
title: "Cross validation"
author: "AnMei Chen"
date: "11/16/2021"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis" , 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## simulate data


```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )
```

look at the data

```{r}
nonlin_df %>% 
  ggplot(aes(x = x , y = y)) +
  geom_point()
```

## cross validation -- by hand

get training and testing datasets

training df sampled 80 of samples from the nonlin_df.
testing df gets 20 samples from the nonlin_df that is not included in the train_df

```{r}

train_df = sample_n(nonlin_df, size = 80)
test_df = anti_join(nonlin_df, train_df, by = "id")

```

Fit three models.

```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = gam( y ~ s(x), data = train_df)
wiggly_mod = gam( y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

can I see what I just did...

```{r}
train_df %>% 
  add_predictions(linear_mod) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() + 
  geom_line(aes( y = pred), color = "red")

train_df %>% 
  add_predictions(smooth_mod) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() + 
  geom_line(aes( y = pred), color = "red")


train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() + 
  geom_line(aes( y = pred), color = "red")
```

smooth model is the best in terms of accuracy.

use gather_predictions() and facet_grid to compare models simultaneously

```{r}
train_df %>% 
  gather_predictions(linear_mod,smooth_mod, wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() + 
  geom_line(aes( y = pred), color = "red") +
  facet_grid(. ~ model)
```

Look at prediction accuracy.

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

